---
title: "Analysis for 'Biases in learned foraging preferences for coloured objects in Trinidadian guppies, Poecilia reticulata'"
output:
  bookdown::html_document2:
    include:
      highlight: pygments
    css: styles.css
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
      smooth_scroll: true
    number_sections: false
    split_by: section
    
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = 'docs/analysis')
  })
---

<div class="topnav">
  <a href="index.html">Guppy colour learning project site</a>
    <a href="https://github.com/wyatt-toure/guppy-colour-learning-project" style = "float: right;">GitHub</a>
    <a class="active" href="analysis.html" style = "float: right;">Analysis</a>
    <a href="methods.html" style = "float: right;">Methods</a>
    <a href="index.html" style = "float: right;">Home</a>
</div>

<p class="author-name">M. Wyatt Toure<span class="affil-mark">1*</span>, Simon M. Reader<span class="affil-mark">1</span></p>

<p class="author-affil"><span class="affil-mark">1</span>McGill University, Department of Biology, 1205 Docteur Penfield, Montreal, Quebec H3A 1B1, Canada</p>
<p><span class="affil-mark">*</span>Corresponding Author: <br>M. Wyatt Toure<br> e-mail: `m_wyatt.toure@mail.mcgill.ca`<br></p>

Date of last update: `r format(Sys.Date(), '%b %d %Y')`

***

## Brief Overview

This page reports the analyses for the experiment described in <span style="text-decoration:underline">'Colour biases in learned foraging preferences in Trinidadian guppies'</span>. The raw data used to conduct these analyses are available in the [colour-learning-project-data.csv](https://github.com/wyatt-toure/guppy-colour-learning-project/blob/main/data/colour-learning-project-data.csv) file. Descriptions of the variables found in the dataset are given in the metadata section of the [README](https://github.com/wyatt-toure/guppy-colour-learning-project#metadata) file. The R script to reproduce the analysis and this site are in the [analysis-script-colour-project.Rmd](https://github.com/wyatt-toure/guppy-colour-learning-project/blob/main/analysis-script-colour-project.Rmd) file. The methods used to produce this data can be found on the [Methods](methods.html) page.

***

```{r library-prep, include=FALSE}
# Loading required packages
library(lme4)
library(tidyr)
library(lmerTest)
library(ggplot2)
library(ggpubr)
library(DHARMa)
library(dplyr)
library(effects)
library(broom)
library(broom.mixed)
library(knitr)
library(emmeans)
library(report)
library(cowplot)
library(tidyext)
library(glmmTMB)
library(MASS)
```

```{r data-prep, include=FALSE}
# Reading in Data 
my.data = read.csv('data/colour-learning-project-data.csv')

# Creating new variables 

## Rewarding object preference
my.data = my.data %>% group_by(id) %>% mutate(rewarding.object.preference = (time.with.trained.object-time.with.untrained.object))

## Green object preference
my.data = my.data %>% mutate(green.object.preference = (case_when(
  rewarding.object.colour == "green" ~ (time.with.trained.object-time.with.untrained.object), 
  rewarding.object.colour == "blue" ~ (time.with.untrained.object-time.with.trained.object) 
)))

# Restrict data to training data
training.data = my.data %>% filter(trial.type =='training')

# Restrict data to only the baseline and re-test data
test.data = my.data %>% filter(trial.type =='test')

# Restrict data to only the baseline data
baseline.data = my.data %>% filter(trial == 0)

# Change trial to factor 
test.data$trial = as.factor(test.data$trial)
baseline.data$trial = as.factor(baseline.data$trial)
```

## Models

We analysed the data from our experiment using linear mixed effect and generalized linear mixed effect models with the `lmer()` and `glmer()` functions from the `lme4` package. P-values and effective degrees of freedom were obtained using the `lmerTest` package. Model residuals were checked they met distributional assumptions with the `DHARMa` package, you can click the 'See Model Residuals' button below the model formulas to see the residual diagnostic plots produced by `DHARMa` for that particular model. 

</br>

### Model 1 -  Preference for the green object at baseline

This first model contains the data for all individual guppies. I looked at the green object preference of all guppies in an intercept only model to see if the green object preference at baseline was significantly different from zero. `green.object.preference` is the time spent near the green object subtracted by the time spent near the blue object

```{r model 1, echo=TRUE}
baseline.data.model = 
  lm(green.object.preference ~ 1,
     data = baseline.data)
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName"> See Model 1 Residuals </button>  
<div id="BlockName" class="collapse"> 

```{r, echo=FALSE}
simulationOutput <- simulateResiduals(fittedModel = baseline.data.model)
plot(simulationOutput)

# Saving plot to figs directory 
ggsave(filename = 'model-1-residual-plot.png', plot = (plot(simulationOutput)), path = 'figs/residual-plots/', device='png', dpi=300)
```

</div>

</br>

##### Result

```{r tidying-model-1, echo=FALSE, message=FALSE}
# Setting table row names
baseline.table.row.name.vec = c("Intercept")
                       
# Converting data frame to tibble
tidy_baseline.model = broom.mixed::tidy(baseline.data.model)

# Changing tibble header names 
tidy_baseline.model = rename(tidy_baseline.model, "Factor" = term, "Estimate" = estimate, "Std. Error" = std.error, "T statistic" = statistic, "P value" = p.value)

# Changing tibble row names 
tidy_baseline.model[1:1,1] <- baseline.table.row.name.vec
```

```{r,  results=TRUE, echo=FALSE}
knitr::kable(tidy_baseline.model%>% mutate_if(is.numeric, round, digits = 3))
```

During the initial test, there was no significant preference for the green object across all guppies (p = `r tidy_baseline.model$'P value' %>%  round(3)`).


```{r baseline-pref-plot, echo=FALSE, message=FALSE, fig.cap="Preference for the green object relative to the blue object across all guppies at baseline. Negative values represent more time spent with the blue object, positive values indicate more time spent with the green object. Data are means +/- 95% CI", fig.id="baseline-pref-plot",  warning=FALSE, message=FALSE}
###### Baseline green object preference plot ######
baseline.data.x.axis.label = 'Initial Test'
ggplot(baseline.data, 
            aes(x = trial, 
                y = green.object.preference)) + theme_classic() + ylab('Green object preference (sec)') + xlab("") + theme(legend.position = "none", axis.text=element_text(size=14), axis.title=element_text(size=14,face="bold"), 
        plot.title = element_text(size=16, hjust=0.5)) + geom_hline(yintercept = 0, linetype = 'dashed', alpha = 0.5) + geom_jitter(width=0.04, alpha = 0.3) + stat_summary(
    geom = 'point',
    fun = 'mean',
    size = 4.5,
    shape = 15) +
  stat_summary(geom = 'errorbar', 
               fun.data = 'mean_ci', position = position_dodge(width=0), width = 0.1)  +
  scale_x_discrete(labels = baseline.data.x.axis.label)

ggsave(filename = 'model-1-baseline-data-plot.png', path = 'figs/', device='png', dpi=300)
```

***

### Model 2 -  Preference for the rewarding object during training

To see whether fish were responsive during training our second model asks whether the preference for the rewarding object changes throughout training and whether the change in rewarding object preference is different between the treatments.  


- **Response variable:** `rewarding.object.preference` is the time (seconds) spent near the rewarding object subtracted by the time spent near the unrewarding object 
- **Fixed effect:** `rewarding.object.colour` is the identity of the rewarding object (blue or green)
- **Fixed effect:** `trial` is the number of the training trial. In this model it is supplied as an integer
- **Random effect:** `id` is the identity of the individual fish 

```{r model 2, echo=TRUE}
training.data.model = 
  lmer(rewarding.object.preference ~ rewarding.object.colour * trial + (1 | id), 
       data = training.data)
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName2"> See Model 2 Residuals </button>  
<div id="BlockName2" class="collapse">  

```{r, echo=FALSE}
simulationOutput <- simulateResiduals(fittedModel = training.data.model, n = 1000)
plot(simulationOutput)

# Saving plot to figs directory 
ggsave(filename = 'model-2-residual-plot.png', plot = (plot(simulationOutput)), path = 'figs/residual-plots/', device='png', dpi=300)
```

There is a slight deviation in the upper quantile but no indication in the residual plot of a gross model misfit.

</div>

```{r tidying-model-2, echo=FALSE, message=FALSE}
# Setting table row names
training.model.table.row.name.vec = c("Intercept", "Reward object colour", "Trial", "Rewarding object colour X Trial")
                       
# Converting data frame to tibble
tidy_training.data.model = broom.mixed::tidy(training.data.model)

# Formatting p value
tidy_training.data.model$p.value = ifelse(tidy_training.data.model$p.value < .001, ## Condition
                      "< .001",  ## if condition is true
                      paste(tidy_training.data.model$p.value %>% round(3))) ## if condition is false

# Getting model confidence intervals
training.data.model.confint = tibble::as_tibble((training.data.model %>% confint()), rownames = "factor")

# Changing tibble header names 
tidy_training.data.model = rename(tidy_training.data.model, "Factor" = term, "Estimate" = estimate, "Std. Error" = std.error, "T statistic" = statistic, "P value" = p.value)

# Changing tibble row names 
tidy_training.data.model[1:4,3] <- training.model.table.row.name.vec
```

</br>

##### Results

```{r,  results=TRUE, echo=FALSE}
knitr::kable(tidy_training.data.model[1:4,] %>% dplyr::select(-group, -effect)%>% mutate_if(is.numeric, round, digits = 3))
```

There was a significant effect of trial. Over the 20 training trials, guppies in the two treatments increased their relative preference for their respective rewarded objects by `r tidy_training.data.model$Estimate[3] %>% round(3)` seconds each trial (Figure \@ref(fig:colour-pref-training-plot), p = `r tidy_training.data.model$'P value'[3]`). There was also a significant effect of rewarded-object colour (p = `r tidy_training.data.model$'P value'[2]`). during training green-rewarded guppies expressed a stronger preference for their rewarded object (the green object) than did blue-rewarded guppies did for the blue object. However, there was no significant interaction effect between rewarding object colour and trial (p = `r tidy_training.data.model$'P value'[4]`), *i.e.*, the rate of increase in object preference over trials did not significantly differ between the treatments.

```{r colour-pref-training-plot, echo=FALSE, message=FALSE, fig.cap="Relative preference for the green object in both treatments during training trials (trials 1-20). Negative values represent more time spent with the blue object, positive values indicate more time spent with the green object. Light lines connect individuals across trials and bold lines represents a linear fit with 95% CI (grey shading). Subjects were consistently rewarded for approaching the blue object (dashed lines) or the green object (solid lines).", fig.id="colour-pref-training-plot",  warning=FALSE, message=FALSE}
###### Time with rewarding object plot during training ######
ggplot(training.data, 
            aes(x = trial, 
                y = green.object.preference, 
                color = rewarding.object.colour,
                linetype = rewarding.object.colour)) + 
  theme_cowplot() + 
  ylab('Green object preference (sec)') + 
  xlab("Trial") + 
  labs(col = "Rewarding object colour") +
  theme(legend.position = "none", 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14,face="bold"), 
        plot.title = element_text(size=16, hjust=0.5)
        ) + 
  scale_color_manual(values=c("#2980b9", "#27ae60")) + 
  scale_linetype_manual(values=c("longdash", "solid")) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_line(aes(group = id), alpha = 0.25) + 
  geom_smooth(method = lm)  

ggsave(filename = 'model-2-colour-pref-training-plot.png', path = 'figs/', device='png', dpi=300)
```

***

 

### Model 3 -  Preference for the rewarded object during testing depending on treatment {#model-3}

For the main effects of training and rewarding object colour on rewarding object preference I fit a generalized linear mixed effects model with a Gaussian distribution which modeled the variances to account for variance heterogeneity using the package [`glmmTMB`](https://cran.r-project.org/web/packages/glmmTMB/index.html). My third model asks whether the preference for the rewarding object changed between baseline and final test and looks for an interaction with rewarded object colour.

- **Response variable:** `rewarding.object.preference` is the time (seconds) spent near the rewarding object subtracted by the time spent near the unrewarding object 
- **Fixed effect:** `rewarding.object.colour` is the identity of the rewarding object (blue or green)
- **Fixed effect:** `trial` is the number of the training trial. In this model it is supplied as a factor where 0 is the baseline trial and 21 is the final trial
- **Random effect:** `id` is the identity of the individual fish 

```{r model 3, echo=TRUE}
test.data.model.glm = 
  glmmTMB(rewarding.object.preference ~  
            trial * rewarding.object.colour + (1|id) + 
    diag(0 + rewarding.object.colour:trial |id), 
  data = test.data, 
  family = gaussian
  )
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName3"> See Model 3 Residuals </button>  
<div id="BlockName3" class="collapse">  

```{r, include=TRUE,echo=FALSE}
simulationOutput <- simulateResiduals(fittedModel = test.data.model.glm, n = 1000)
plot(simulationOutput)

# Saving plot to figs directory 
ggsave(filename = 'model-3-residual-plot.png', plot = (plot(simulationOutput)), path = 'figs/residual-plots/', device='png', dpi=300)
```

</div>



```{r tidying-model-3, echo=FALSE, message=FALSE}
# Setting table row names
test.model.table.row.name.vec = c("Intercept", "Trial", "Rewarding object colour", "Rewarding object colour X Trial")
                       
# Converting data frame to tibble
tidy_test.data.model = broom.mixed::tidy(test.data.model.glm)

# Formatting p value
tidy_test.data.model$p.value = ifelse(tidy_test.data.model$p.value < .001, ## Condition
                      "< .001",  ## if condition is true
                      paste(tidy_test.data.model$p.value %>% round(3))) ## if condition is false

# Getting model confidence intervals
# test.data.model.confint = tibble::as_tibble((test.data.model %>% confint()), rownames = "factor")

# Changing tibble header names 
tidy_test.data.model = rename(tidy_test.data.model, "Factor" = term, "Estimate" = estimate, "Std. Error" = std.error, "T statistic" = statistic, "P value" = p.value)

# Changing tibble row names 
tidy_test.data.model[1:4,4] <- test.model.table.row.name.vec
```

</br>

##### Results

```{r, results=TRUE, echo=FALSE}
knitr::kable(tidy_test.data.model[1:4,] %>% dplyr::select(-group, -effect, -component) %>% mutate_if(is.numeric, round, digits = 3), caption = "Summary table for Model 3. Estimates ± standard error (SE) of the effects of trial and rewarding object colour on the rewarding object preference from the generalized linear mixed effect model containing the effects Trial, Rewarding object colour, and their interaction effect (Trial X Rewarding object colour).")
```

When comparing the initial and final preference test, both conducted after training and without food rewards present, we found a significant interaction effect between test and rewarding object colour (p = `r tidy_test.data.model$'P value'[4]`). Guppies that had been green-rewarded had a shift in their rewarded object preference that was on average `r tidy_test.data.model$Estimate[4] %>% round(1)` seconds stronger than the shift in rewarded object preference of guppies trained to blue (Figure \@ref(fig:test-data-pref-plot)). These results were unaffected by the removal of one fish that did not feed during training ([See ESM model 2](#esm-model-2)).

```{r, include=FALSE}
# We see that there is a significant interaction effect. To see whether both blue-trained and green-trained guppies significantly shift their rewarded object preference

# in response to training I will use the package emmeans to conduct pairwise comparisons of baseline and re-test data across treatments

# Is the difference in rewarding object preference between trials (baseline and re-test) significant in both object treatments (blue-trained and green-trained)?

test.data.model.pairwise.comparisons = emmeans(test.data.model.glm, specs = ~ trial:rewarding.object.colour)

test.data.model.pairwise.comparisons

blue0 = c(1,0,0,0)
blue21 = c(0,1,0,0)
green0 = c(0,0,1,0)
green21 = c(0,0,0,1)



custom.contrasts = contrast(test.data.model.pairwise.comparisons, 
                            method = list("21 blue - 0 blue" = blue21- blue0,
                                          "21 green - 0 green " = green21 - green0,
                                          "21 green - 21 blue" = green21 - blue21,
                                          "0 green - 0 blue" = green0- blue0), adjust = "mvt") %>% summary(infer = TRUE)

# Pull out the summary statistics of the contrasts and turn them into a table
pairwise.comparison.means.table = as.data.frame(test.data.model.pairwise.comparisons)

pairwise.comparison.contrasts.table = as.data.frame(custom.contrasts)

# Rounding contrast table values

formatted_p.pairwise.comparison.contrasts.table = ifelse(pairwise.comparison.contrasts.table$p.value < .001, ## Condition
                      "< .001",  ## if condition is true
                      paste("=", pairwise.comparison.contrasts.table$p.value %>% round(3))) ## if condition is false

pairwise.comparison.contrasts.table$estimate = pairwise.comparison.contrasts.table$estimate %>% round(3)
pairwise.comparison.contrasts.table$lower.CL = pairwise.comparison.contrasts.table$lower.CL %>% round(3)
pairwise.comparison.contrasts.table$upper.CL = pairwise.comparison.contrasts.table$upper.CL %>% round(3)
pairwise.comparison.contrasts.table$df = pairwise.comparison.contrasts.table$df %>% round(3)
```


```{r test-data-pref-plot, echo=FALSE, message=FALSE, fig.cap="Changes in object preference from an initial test before training to a final test after training. During training, fish were rewarded for approaching the blue object (blue squares and lines) or the green object (green squares and lines). At test, no food reward was present. Dashed line represents an equal preference for either object. Data are means ± 95% CI; lighter points and lines are data for each individual. The final preference for green-rewarded fish is stronger than that of fish that has been rewarded for approaching the blue object, but both sets of subjects increased their preference for the rewarded object compared to the initial test preference.", fig.id="test-data-pref-plot", warning=FALSE, message=FALSE}
###### Time with rewarding object plot during testing ######
testing.data.x.axis.labels = c("Initial test", "Final test")
object.labels = c("Blue-trained", "Green-trained")
names(object.labels) <- c("blue", "green")

ggplot(test.data, 
            aes(x = trial, 
                y = rewarding.object.preference, 
                color = rewarding.object.colour,
                shape = rewarding.object.colour)) + 
  theme_cowplot() + 
  geom_jitter(width=0, alpha = 0.3) + 
  stat_summary(geom = 'point', fun = 'mean', size = 4.5) +
  stat_summary(geom = 'errorbar', fun.data = 'mean_ci', position = position_dodge(width=0), width = 0.1) + 
  ylab('Rewarding object preference (sec)') + 
  xlab("Trial") + 
  labs(col = "Rewarding object colour") +
  theme(legend.position = "none", 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14,face="bold"), 
        plot.title = element_text(size=16, hjust=0.5)
        ) + 
  scale_color_manual(values=c("#2980b9", "#27ae60")) + 
  scale_shape_manual(values=c(15,16)) +
  geom_hline(yintercept = 0, linetype = 'dashed', alpha = 0.6) +
  geom_line(aes(group = id), alpha = 0.2) +
  scale_x_discrete(labels = testing.data.x.axis.labels) + 
  facet_grid(~rewarding.object.colour, labeller = labeller(rewarding.object.colour = object.labels)) +
  stat_summary(fun=mean, geom="line", aes(group = rewarding.object.colour))

ggsave(filename = 'model-3-test-data-pref-plot.png', path = 'figs/', device='png', dpi=300)
```



</br>

#### Post-hoc Comparisons

To determine whether the means of the final rewarding object preference for the two treatments were different I conducted post-hoc comparisons with the package [`emmeans`](https://cran.r-project.org/package=emmeans). I compared the following means:

- Final test blue-trained and initial test blue-trained
- Final test green-trained and initial test green-trained
- Final test green-trained and final test blue-trained
- Initial test green-trained and initial test blue-trained

```{r, echo=TRUE, eval=FALSE}
test.data.model.pairwise.comparisons = 
  emmeans(test.data.model.glm, specs = ~ trial:rewarding.object.colour)

custom.contrasts = contrast(test.data.model.pairwise.comparisons, 
                            method = list("21 blue - 0 blue" = blue21- blue0,
                                          "21 green - 0 green " = green21 - green0,
                                          "21 green - 21 blue" = green21 - blue21,
                                          "0 green - 0 blue" = green0- blue0), 
                            adjust = "mvt") %>% summary(infer = TRUE)
```

</br>

##### Results

```{r contrasts-table, echo=FALSE, fig.id="contrasts-table", warning=FALSE}
pairwise.comparison.contrasts.table.show = pairwise.comparison.contrasts.table

pairwise.comparison.contrasts.table.show$p.value = 
  case_when(pairwise.comparison.contrasts.table.show$p.value < .001 ~ paste('< .001'),
            (pairwise.comparison.contrasts.table.show$p.value < .05 & pairwise.comparison.contrasts.table.show$p.value >= .001) ~ paste(as.character(pairwise.comparison.contrasts.table.show$p.value %>% round(3))),
            pairwise.comparison.contrasts.table.show$p.value >= .05 ~ paste(pairwise.comparison.contrasts.table.show$p.value %>% round(3)))

knitr::kable(pairwise.comparison.contrasts.table.show %>% dplyr::select(contrast, estimate,lower.CL,upper.CL, df,p.value), caption = "Table of post-hoc tests with a multivariate-t adjustment for multiple comparisons of a selected set of means. The numbers represent the initial test (‘Initial’) and the final test (‘Final’). The colour corresponds to the identity of the rewarding object (blue for blue-rewarded guppies, green for green-rewarded guppies). Values are all rounded to 3 decimal places. CL = confidence limit.")

```

Post-hoc comparisons (Table \@ref(tab:contrasts-table)) reveal that initially, before training, there was no significant difference in the strength of preference for the rewarded object between the treatments (p `r formatted_p.pairwise.comparison.contrasts.table[4]`). The shift in rewarded object preference between the initial and final preference tests was significant for green-trained guppies but not for blue-trained guppies: green trained guppies increased their preference for the green object by `r abs(pairwise.comparison.contrasts.table$estimate[2] %>% round(1))` seconds from initial to final (p `formatted_p.pairwise.comparison.contrasts.table[1]`) test whereas blue-trained guppies increased their preference for the blue object by `r abs(pairwise.comparison.contrasts.table$estimate[1] %>% round(1))` seconds an effect that was not statistically significant (p `r formatted_p.pairwise.comparison.contrasts.table[1]`). At final test, green-rewarded guppies had a significantly stronger preference for the previously rewarded object compared to the blue-rewarded guppies.

***


```{r, include = FALSE}
#### Get feeding data #####
# Group by ID and count the number of sessions in which an individual ate 
feeding = my.data %>% group_by(id) %>%  count(feeding.count = ate == 'yes')
feeding
# Remove NAs from this 
feeding = na.omit(feeding)
feeding
# Count only the yeses
feeding = feeding %>% filter(feeding.count == 'TRUE')
# Remove the column session.ate to keep only the counts
feeding = feeding %>% dplyr::select(-feeding.count)
feeding

# Add the feeding values to the main data frame and replace NAs with 0 
my.feeding.data = left_join(baseline.data, feeding, by ="id")
my.feeding.data = my.feeding.data %>% replace_na(list(n = 0))
my.feeding.data
my.feeding.data = rename(my.feeding.data, feeding.count = n)
my.feeding.data =  my.feeding.data %>% dplyr::select(id,feeding.count,rewarding.object.colour)
my.feeding.data
```

### Model 4 - Is there a difference in feeding attempts between treatments? {#model-4}

A discrepancy in reinforcement between treatments may influence performance on a final preference test. To see whether there was a difference in feeding between treatments I counted the number of trials in which an individual fish ate throughout all of training and compared the feeding counts between treatments. To do this I fit a generalized linear model with a negative binomial distribution. The response variable 'feeding count' is a sum of the number of trials in which a guppy ate.

- **Response variable:** `feeding.count` is the number of trials in which an individual fish ate
- **Fixed effect:** `rewarding.object.colour` is the identity of the rewarding object (blue or green)

##### Model

```{r model 4, echo=TRUE}
feeding.data.model = 
  glm.nb(feeding.count ~ rewarding.object.colour, 
      data = my.feeding.data)
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName4"> See Model 4 Residuals </button>  
<div id="BlockName4" class="collapse">  

```{r, echo=FALSE}
simulationOutput <- simulateResiduals(fittedModel = feeding.data.model)
plot(simulationOutput)

# Saving plot to figs directory 
ggsave(filename = 'model-4-residual-plot.png', plot = (plot(simulationOutput)), path = 'figs/residual-plots/', device='png', dpi=300)
```

</div>

```{r tidying-model-4, echo=FALSE, message=FALSE}
# Setting table row names
feeding.model.table.row.name.vec = c("Intercept", "Rewarding object colour")
                       
# Converting data frame to tibble
tidy_feeding.data.model = broom.mixed::tidy(feeding.data.model)

# Getting model confidence intervals
feeding.data.model.confint = tibble::as_tibble((feeding.data.model %>% confint()), rownames = "factor")

# Changing tibble header names 
tidy_feeding.data.model = rename(tidy_feeding.data.model, "Factor" = term, "Estimate" = estimate, "Std. Error" = std.error, "T statistic" = statistic, "P value" = p.value)

# Changing tibble row names 
tidy_feeding.data.model[1:2,1] <- feeding.model.table.row.name.vec
```

</br>

##### Results

```{r, results=TRUE, echo=FALSE}
knitr::kable(tidy_feeding.data.model[1:2,]%>% mutate_if(is.numeric, round, digits = 3))
```

We found no significant difference in the number of trials individuals fed between green-rewarded and blue-rewarded fish (Figure \@ref(fig:feeding-count-plot), p = `r tidy_feeding.data.model$'P value'[2] %>% round(3)`). Removing the one individual that did not feed during training gave a similar result (p = 0.87). We also incorporated feeding count as a covariate in [Model 3](#model-3), finding the same pattern of results (See [ESM Model 2](#esm-model-2)).

```{r feeding-count-plot, echo=FALSE, message=FALSE, fig.cap="Average number of trials in which a fish fed during training. Data are means ± 95% confidence intervals with probability density functions of the data (showing the negative binomial distribution) to the right of the raw data.", fig.id="training-data-ate-plot", warning=FALSE}
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
ggplot(my.feeding.data,
             aes(x=rewarding.object.colour,
                 y=feeding.count, 
                 fill = rewarding.object.colour, 
                 colour = rewarding.object.colour)) +
  geom_point(position = position_jitter(width = 0.05), alpha = 0.8) +
  geom_flat_violin(aes(fill = rewarding.object.colour),position = position_nudge(x = .25, y = 0),adjust = 0.7, alpha = 0.4, trim = FALSE, color=NA) + 
  stat_summary(geom = 'point', fun = 'mean', size = 4.5, shape = 15) +
  stat_summary(geom = 'errorbar', fun.data = 'mean_ci', position = position_dodge(width=0), width = 0.1) + ylim(-3,23) +
  ylab('Rewarding object preference')+xlab('Trial') + theme_cowplot()+ guides(fill = FALSE, colour = FALSE) + ylab('Number of trials fed') + xlab("Rewarding object colour") + labs(col = "Rewarding object colour") +
  theme(legend.position = "top", axis.text=element_text(size=14), axis.title=element_text(size=14,face="bold"), 
        plot.title = element_text(size=16, hjust=0.5)) +
  scale_color_manual(values=c("#2980b9", "#27ae60"))+
  scale_fill_manual(values=c("#2980b9", "#27ae60"))

ggsave(filename = 'model-4-feeding-count-plot.png', path = 'figs/', device='png', dpi=300)
```

***


## ESM Models

In this section we describe models not included in the main text.


### Addressing feeding count

The models described in this section were run to address the potential role of feeding count on test performance. The concern was that the results may be explained by differential levels of reinforcement between the groups or between individuals during training. To ensure our results were robust to matters arising from feeding count variation we: 

1) Removed an individual that did not feed and re-ran the analysis in [Model 3](#model-3)
2) Included feeding count as a co-variate and re-ran the analysis in [Model 3](#model-3)

</br>

#### ESM Model 1 - Removing individual that did not feed

```{r, include=FALSE}
test.feeding.data = left_join(test.data, feeding, by ="id")
test.feeding.data = test.feeding.data %>% replace_na(list(n = 0))
test.feeding.data = rename(test.feeding.data, feeding.count = n)
retest.feeding.data = test.feeding.data %>% filter(trial == "21")
retest.feeding.data = retest.feeding.data %>% mutate(time.with.both.objects = time.with.trained.object+time.with.untrained.object)
```

There is a fish that did not eat during any trials, however removing this individual does not change the conclusions 

```{r, echo=TRUE}
test.feeding.data.low.feeders.removed = 
  test.feeding.data %>% filter(feeding.count > 0)
```


##### Model

```{r, echo=TRUE}
test.feeding.data.low.feeders.removed.model = 
  glmmTMB(rewarding.object.preference ~  
            trial * rewarding.object.colour + (1|id) + 
    diag(0 + rewarding.object.colour:trial |id),
    data = test.feeding.data.low.feeders.removed, 
    family = gaussian
    )
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName6"> See ESM Model 1 Residuals </button>  
<div id="BlockName6" class="collapse">  

```{r, echo=FALSE}
simulationOutput <- simulateResiduals(fittedModel = test.feeding.data.low.feeders.removed.model)
plot(simulationOutput)

# Saving plot to figs directory 
ggsave(filename = 'ESM-model-1-residual-plot.png', plot = (plot(simulationOutput)), path = 'figs/residual-plots/', device='png', dpi=300)
```

</div>

</br>

```{r, echo=FALSE, results=FALSE}
# Setting table row names
feeding.data.low.feeders.removed.model.table.row.name.vec = c("Intercept", "Trial", "Rewarding object colour", "Rewarding object colour X Trial")

tidy_fit.test.feeding.data.low.feeders.removed = broom.mixed::tidy(test.feeding.data.low.feeders.removed.model)

# format p value 
tidy_fit.test.feeding.data.low.feeders.removed$p.value = ifelse(tidy_fit.test.feeding.data.low.feeders.removed$p.value < .001, ## Condition
                      "< .001",  ## if condition is true
                    paste(tidy_fit.test.feeding.data.low.feeders.removed$p.value %>% round(3))) ## if condition is false

# Changing tibble header names 
tidy_fit.test.feeding.data.low.feeders.removed = rename(tidy_fit.test.feeding.data.low.feeders.removed, "Factor" = term, "Estimate" = estimate, "Std. Error" = std.error, "T statistic" = statistic, "P value" = p.value)

tidy_fit.test.feeding.data.low.feeders.removed[1:4,4] <- feeding.data.low.feeders.removed.model.table.row.name.vec
```

##### Results

```{r, results=TRUE, echo=FALSE}
knitr::kable(tidy_fit.test.feeding.data.low.feeders.removed[1:4,] %>% dplyr::select(-group, -effect, -component)%>% mutate_if(is.numeric, round, digits = 3))
```

***

#### ESM Model 2 - Including feeding count as a covariate {#esm-model-2}

With this next model we looked to see whether including the amount of trials an individual fed as a covariate in the model changes the conclusions. 

- **Response variable:** `rewarding.object.preference` is the time (seconds) spent near the rewarding object subtracted by the time spent near the unrewarding object 
- **Fixed effect:** `rewarding.object.colour` is the identity of the rewarding object (blue or green)
- **Fixed effect:** `trial` is the number of the training trial. In this model it is supplied as an integer
- **Random effect:** `id` is the identity of the individual fish 
- **Covariate:** `feeding.count` is the number of trials in which an individual fish ate

```{r, echo = TRUE}
test.data.feeding.controlled.model = 
  glmmTMB(rewarding.object.preference ~  
            trial * rewarding.object.colour + feeding.count + (1|id) + 
    diag(0 + rewarding.object.colour*trial |id), 
  data = test.feeding.data, family = gaussian)
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName7"> See ESM Model 2 Residuals </button>  
<div id="BlockName7" class="collapse">  

```{r, echo=FALSE}
simulationOutput <- simulateResiduals(fittedModel = test.data.feeding.controlled.model)
plot(simulationOutput)

# Saving plot to figs directory 
ggsave(filename = 'ESM-model-2-residual-plot.png', plot = (plot(simulationOutput)), path = 'figs/residual-plots/', device='png', dpi=300)
```

</div>

</br>
```{r, echo=FALSE, results=FALSE}
# Setting table row names
test.data.feeding.controlled.model.table.row.name.vec = c("Intercept", "Trial", "Rewarding object colour", "Feeding count", "Rewarding object colour X Trial")

tidy_fit.test.data.feeding.controlled.model = broom.mixed::tidy(test.data.feeding.controlled.model)

# formatted p value 
tidy_fit.test.data.feeding.controlled.model$p.value = ifelse(tidy_fit.test.data.feeding.controlled.model$p.value < .001, ## Condition
                      "< .001",  ## if condition is true
                      paste(tidy_fit.test.data.feeding.controlled.model$p.value %>% round(3))) ## if condition is false

# Changing tibble header names 
tidy_fit.test.data.feeding.controlled.model = rename(tidy_fit.test.data.feeding.controlled.model, "Factor" = term, "Estimate" = estimate, "Std. Error" = std.error, "T statistic" = statistic, "P value" = p.value)

tidy_fit.test.data.feeding.controlled.model[1:5,4] <- test.data.feeding.controlled.model.table.row.name.vec
```

##### Results

```{r, results=TRUE, echo=FALSE}
knitr::kable(tidy_fit.test.data.feeding.controlled.model[1:5,]%>% dplyr::select(-group, -effect, -component)%>% mutate_if(is.numeric, round, digits = 3), caption = "Summary table for a modification of Model 3. This model is the same as that described in Table S1 except it includes feeding count as a covariate. Estimates ± standard error (SE) of the effects of trial and rewarding object colour of the rewarding object colour from the generalized linear mixed effect model containing the effects (Trial, Rewarding object colour, and their interaction effect).")
```


The main results do not change if I control for feeding count. The above table is the output feeding controlled model. Below we have the output table for the non-feeding-count controlled model from [Model 3](#model-3).

```{r, results=TRUE, echo=FALSE}
knitr::kable(tidy_test.data.model[1:4,] %>% dplyr::select(-group, -effect, -component) %>% mutate_if(is.numeric, round, digits = 3))
```

In both models the p-values are similar. While the effect of feeding count is not significant (p = `r tidy_fit.test.data.feeding.controlled.model$'P value'[4]`) the effect of feeding count trends in the expected direction in our feeding count controlled model. As feeding count increases the preference for the rewarding object colour also increases.

***

### Feeding and object preferences

This is a model I (Wyatt) ran just because I was curious. 

#### ESM Model 3 - Preference for the rewarded object during training based on feeding

This model asks whether the time spent near the rewarding object during a training session is influenced by whether a fish ate or not. Here trial is a variable containing the training trials (1-20). It is supplied as a random effect along with individual ID in the model. 

- **Responsible variable:** `rewarding.object.preference` is the time (seconds) spent near the rewarding object subtracted by the time spent near the unrewarding object 
- **Fixed effect:** `ate` is a binary factor (Yes or No) and corresponds to whether the fish ate in a given trial
- **Fixed effect:** `trial` is the number of the training trial. In this model it is supplied as an integer
- **Random effect:** `id` is the identity of the individual fish 

##### Model

```{r esm-model-1, echo=TRUE}
training.data.model.rewarding.object = 
  lmer(rewarding.object.preference ~ ate + (1 | id) + (1 | trial),
       data = training.data)
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName5"> See ESM Model 3 Residuals </button>  
<div id="BlockName5" class="collapse">  

```{r, echo=FALSE}
simulationOutput <- simulateResiduals(fittedModel = training.data.model.rewarding.object, n = 1000)
plot(simulationOutput)

# Saving plot to figs directory 
ggsave(filename = 'ESM-model-3-residual-plot.png', plot = (plot(simulationOutput)), path = 'figs/residual-plots/', device='png', dpi=300)
```

There is a significant deviation from uniformity as indicated by the significant Kolmogorov-Smirnov test. However, this model has a particularly large sample size (n = 456) so even slight deviations will be significant. Looking at the effect size of the deviation (D = `r testUniformity(simulationOutput)[1]`) shows that it is minor (D < 0.1) and visual inspection does not suggest large deviations in the residuals so our model is still appropriate.

</div>

</br>

```{r, echo=FALSE, results=FALSE}
# Setting table row names
training.data.model.rewarding.object.table.row.name.vec = c("Intercept", "Ate")

# Summary of esm model 3
tidy_fit.rewarding.object = broom.mixed::tidy(training.data.model.rewarding.object)
tidy_fit.rewarding.object$p.value = ifelse(tidy_fit.rewarding.object$p.value < .001, ## Condition
                      "< .001",  ## if condition is true
                      paste(tidy_fit.rewarding.object$p.value %>% round(3))) ## if condition is false

# Changing tibble header names 
tidy_fit.rewarding.object = rename(tidy_fit.rewarding.object, "Factor" = term, "Estimate" = estimate, "Std. Error" = std.error, "T statistic" = statistic, "P value" = p.value)

tidy_fit.rewarding.object[1:2,3] <- training.data.model.rewarding.object.table.row.name.vec
```

##### Results

```{r, results=TRUE, echo=FALSE}
knitr::kable(tidy_fit.rewarding.object[1:2,] %>% dplyr::select(-group, -effect)%>% mutate_if(is.numeric, round, digits = 3))
```


```{r, include=FALSE}
# Summary of model 4

summary(training.data.model.rewarding.object)

# This is needed to get the model output in a format that can be referenced later

tidy_fit.rewarding.object = broom.mixed::tidy(training.data.model.rewarding.object)

# This formats the p-value depending on whether it is less than 0.001. If that is true
# it will just report p < .001 otherwise it will give the exact value rounded to 3 decimal points

formatted_p.rewarding.object = ifelse(tidy_fit.rewarding.object$p.value < .001, ## Condition
                      "< .001",  ## if condition is true
                      paste("=", tidy_fit.rewarding.object$p.value %>% round(3))) ## if condition is false
```

Throughout all of training, fish that ate during a session spent on average `r tidy_fit.rewarding.object$estimate[2] %>% round(1)` more seconds near the rewarded object compared to fish that did not eat (Figure, \@ref(fig:training-data-ate-plot), p `r formatted_p.rewarding.object[2]`). Fish that spent more trials eating may therefore have received more reinforcement for the object-food association. This result is what led me to investigate whether there was a difference in food reinforcement between the two treatments in [Model 4](#model-4). 
</br>

```{r training-data-ate-plot, echo=FALSE, message=FALSE, fig.cap="Preference for the rewarding object during training based on whether an individual ate during a trial or not. Dashed line represents the no preference value. In panel A, Data are means ± 95% CI. In panel B, the data consist of all the feeding choices of all individuals across all training trials. Data are means ± 95% CI with probability density functions to the right of the raw data.", fig.id="training-data-ate-plot", warning=FALSE}
###### Time with rewarding object plot during training [gaussian] ######
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")

p1 = ggplot(training.data, 
            aes(x = trial, 
                y = rewarding.object.preference, 
                color = ate)) + theme_cowplot() + geom_jitter(width=0.05, alpha = 0.2) + stat_summary(
    geom = 'point',
    fun = 'mean',
    size = 4.5,
    shape = 15) +
  stat_summary(geom = 'errorbar', 
               fun.data = 'mean_ci', position = position_dodge(width=0), width = 0.1) + ylab('Rewarding object preference (sec)') + xlab("Trial") + labs(col = "Did this individual eat this trial?") +
  theme(legend.position = "top", axis.text=element_text(size=14), axis.title=element_text(size=14,face="bold"), 
        plot.title = element_text(size=16, hjust=0.5), panel.border = element_rect(colour = "black", fill=NA, size=1), axis.line.x = element_blank(), axis.line.y = element_blank()) + scale_color_manual(values=c('#696667','#0f85a0')) + 
  geom_hline(yintercept = 0, linetype = 'dashed') + ylim(-310,310)

p2 = ggplot(training.data, 
            aes(x = trial.type, 
                y = rewarding.object.preference, 
                color = ate)) + theme_cowplot() + geom_jitter(width=0.05, alpha = 0.1) +
  geom_flat_violin(aes(fill = ate),position = position_nudge(x = .25, y = 0),adjust = 1, alpha = 0.6, trim = TRUE, color=NA) + stat_summary(
    geom = 'point',
    fun = 'mean',
    size = 3,
    shape = 15) +
  stat_summary(geom = 'errorbar', 
               fun.data = 'mean_ci', position = position_dodge(width=0), width = 0.15) + xlab("") + ylab("Rewarding object preference (sec)")+
  theme(legend.position = "none", axis.text=element_text(size=14), axis.title=element_text(size=12.5,face="bold"), 
        plot.title = element_text(size=16, hjust=0.5), axis.ticks.x = element_blank(), axis.line.y = element_blank(), axis.line.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y =  element_blank(), axis.title.y = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_color_manual(values=c('#696667','#0f85a0')) + scale_fill_manual(values=c('#696667','#0f85a0')) + 
  geom_hline(yintercept = 0, linetype = 'dashed') + ylim(-310,310) +
  scale_x_discrete(labels = "")

ggarrange(p1,p2, widths = c(1.5,1),
          labels = "AUTO",
          label.x = c(0.15,0),
          hjust = -2,
          vjust = 3,
          common.legend = TRUE)

ggsave(filename = 'esm-model-3-training-data-ate-plot.png', path = 'figs/', device='png', dpi=300)
```


***

</br>

## Tools used and References

A complete list of the tools used is produced below


```{r, echo=FALSE, warning=FALSE}
knitr::kable(as.data.frame(report(sessionInfo())))
```

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
