---
title: "Datasheet for colour learning part 2"
output:
  bookdown::html_document2:
    include:
      highlight: pygments
    number_sections: false
    split_by: section
    
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = 'docs/datasheet-part2')
  })
---

```{r library-prep, include=FALSE}
# Loading required packages
library(dplyr)
library(randomizr)
```

## Assigning treatments

To assign individuals to treatments I used the complete_ra() function from the
randomizr package to randomly assign equal amounts of n = 16 individuals to
either blue or green trained treatments.

Next I used rbinom to return a random sequence of 1s and 0s which I converted
to left and right which would determine which side of the tank an object would
end up on. I used a random seed which produced equal amounts of 1s and 0s.

The I run a series of dplyr calls which creates a new variable called trial
that counts from 1 to nrow() of the data frame separated by id. There are 20
levels of ID.

```{r treatment-assignment, echo=FALSE}
# Creating IDs
id <- as.factor(c(
  "1a", "1b", "2a", "2b", "3a", "3b", "4a", "4b",
  "5a", "5b", "6a", "6b", "7a", "7b", "8a", "8b"
))

# Get treatment assignments
set.seed(123)
treatment <- complete_ra(
  N = 16,
  num_arms = 2,
  conditions = c("green-trained", "blue-trained")
)

# Get object sides for training
set.seed(52)
object_sides_training <- rbinom(20, 1, 0.5)

# Create a data frame with trials and object sides
data_sheet <- data.frame(id = rep(id, times = 20)) %>%
  group_by(id) %>%
  mutate(trial = 1:n()) %>%
  ungroup() %>%
  mutate(
    object.side =
      ifelse(
        trial %% 2 == 1,
        object_sides_training, 1 - object_sides_training
      )
  )

# Change object side to factor and re-code from numbers to words
data_sheet$object.side <- as.factor(data_sheet$object.side)
data_sheet$object.side <- recode_factor(data_sheet$object.side,
  "1" = "left",
  "0" = "right"
)

# add individual treatments to data sheet
data_sheet <- cbind(data_sheet, treatment)
```



