---
title: " "
output:
  bookdown::html_document2:
    include:
      highlight: pygments
    number_sections: false
    split_by: section
    
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = 'docs/datasheet-part2')
  })
---

```{r library-prep, include=FALSE}
# Loading required packages
library(dplyr)
library(randomizr)
```

## Assigning treatments

To assign individuals to treatments I used the complete_ra() function from the
randomizr package to randomly assign equal amounts of n = 16 individuals to
either blue or green trained treatments. I picked an arbitrary random seed.

Next I used rbinom to return a random sequence of 1s and 0s which I converted to
left and right which would determine which side of the tank an object would end
up on. To get the random seed I randomly picked numbers until I got one which
produced equal amounts of 1s and 0s.

The I run a series of dplyr calls which creates a new variable called trial that
counts from 1 to nrow() of the data frame separated by id. There are 20 levels
of ID.

```{r treatment-assignment, echo=FALSE}
# Creating IDs
id <- as.factor(c(
  "1a", "1b", "2a", "2b", "3a", "3b", "4a", "4b",
  "5a", "5b", "6a", "6b", "7a", "7b", "8a", "8b"
))

# Get treatment assignments
set.seed(123)
treatment <- complete_ra(
  N = 16,
  num_arms = 2,
  conditions = c("green-trained", "blue-trained")
)

set.seed(123)
object_pair <- complete_ra(
  N = 20,
  num_arms = 4,
  conditions = c("pair-1", "pair-2", "pair-3", "pair-4")
)

# Get object sides for training
set.seed(52)
object_sides_training <- rbinom(20, 1, 0.5)

# Create a data frame with trials and object sides
data_sheet <- data.frame(id = rep(id, times = 20)) %>%
  group_by(id) %>%
  mutate(trial = 1:n()) %>%
  ungroup() %>%
  mutate(
    object.side =
      ifelse(
        trial %% 2 == 1,
        object_sides_training, 1 - object_sides_training
      )
  )

# Change object side to factor and re-code from numbers to words
data_sheet$object.side <- as.factor(data_sheet$object.side)
data_sheet$object.side <- recode_factor(data_sheet$object.side,
  "1" = "left",
  "0" = "right"
)

# Assigning object pairs to trials
data_sheet$trial <- as.factor(data_sheet$trial) # convert trial to factor
names(object_pair) <- levels(data_sheet$trial) # get the list of trial numbers to associate object pairs to
data_sheet$object_pair <- object_pair[as.character(data_sheet$trial)]

# Add individual treatments to data sheet
data_sheet <- cbind(data_sheet, treatment)
```

```{r, echo=FALSE}
data_sheet %>% knitr::kable()
```




```{r library-prep2, include=FALSE}
library(ggplot2)
library(scales)
library(cowplot)
```

```{r data-prep, include=FALSE}
trial.type = c(
  "Initial Test", "Training","Training",
"Training", "Training", "Training", "Training", "Training","Training","Training",
"Training","Training","Training","Training","Training","Training","Training","Training",
"Training", "Training", "Training", "Final Test", "Refresher", "Refresher", "Generalization 1","Generalization 2")

trial = c(0, 1,2,3,4,5,6,7,8,9,10,
          11,12,13,14,15,16,17,18,19,20,
          21,22,23,24,25)

trial.state = c(
  "Initial","Initial","Initial",
"Initial","Initial","Initial","Initial",
"Training","Training","Training","Training","Training",
"Training","Training","Training","Training","Training",
"Training","Training","Training","Training","Final Test",
"Refresher","Refresher","Generalization 1","Generalization 2"
)

# Tableau colour palettes
blue = "#4e79a7"
green = "#59a14f"
brown = "#9c755f"
orange = "#f28e2b"
yellow = "#edc948"
grey = "#bab0ac"
red = "#e15759"
purple = "#b07aa1"
cyan = "#76b7b2"
pink = "#ff9da7"
```

```{r data-prep2, include=FALSE}
timeline.df = data.frame(trial.type, trial, trial.state)

trial.type.colors <- c(orange, cyan, blue, orange,  red,  green)
trial.type.shapes <- c(15, 2, 1, 0, 16, 19)
```

```{r data-prep3, include=FALSE}
positions <- c(0.5, 0.5, 0.5, 0.5, 0.5, 0.5)
directions <- c(1, 1)

line_pos <- data.frame(
    "trial"=unique(timeline.df$trial),
    "position"=rep(positions, length.out=length(unique(timeline.df$trial))),
    "direction"=rep(directions, length.out=length(unique(timeline.df$trial)))
)

timeline.df <- merge(x=timeline.df, y=line_pos, by="trial", all = TRUE)
timeline.df <- timeline.df[with(timeline.df, order(trial, trial.type)), ]

text_offset <- 0.1

timeline.df$trial_count <- ave(timeline.df$trial==timeline.df$trial, timeline.df$trial, FUN=cumsum)
timeline.df$text_position <- (timeline.df$trial_count * text_offset * timeline.df$direction) + timeline.df$position

head(timeline.df)
```

```{r, include=FALSE}
# Initial ggplot version of timeline commented out. If you see this Hi! :) 
ggplot(timeline.df,aes(x=trial,
                       y=0,
                       color=trial.type,
                       shape = trial.type,
                       label = trial)
       ) +
  scale_color_manual(values=trial.type.colors, drop = FALSE) +
  scale_shape_manual(values = trial.type.shapes) +
  theme_cowplot() +
  geom_hline(yintercept=0, color = "white", size=0.3) +
  geom_point(aes(y=0), size=4) +
  theme(
    axis.line.y=element_blank(),
    axis.text.y=element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank(),
    axis.ticks.y=element_blank(),
    axis.text.x =element_blank(),
    axis.ticks.x =element_blank(),
    axis.line.x =element_blank(),
    plot.title = element_text(hjust = 0.5,  vjust = -30),
    legend.position = "none"
                ) +
  guides(colour = guide_legend(nrow = 2)) +
  geom_text(aes(label=trial),hjust=.7, vjust=-2,size=3, fontface = "bold") +
  labs(color = "Trial Type", shape = "Trial Type")
```




