---
title: "Assigning treatments and object sides"
author: "M. Wyatt Toure"
---

```{r library-prep, include=FALSE}
# Loading required packages
library(dplyr)
library(randomizr)
```

## Summary

For experiment 2 we performed treatment and object side assignments
programmatically in R. This document shows the code we used to make these
assignments. For object sides some of the side assignments were derived from
previous, programmatic assignments (*e.g.*, the next trial was just the previous
trial reversed to counter-balance across test trials). I have specified when
this is the case and for which trials. 

## Assigning treatments

To assign individuals to treatments I used the `complete_ra()` function from the
`randomizr` package to randomly assign equal amounts of n = 16 individuals to
either blue or green trained treatments. I picked an arbitrary random seed.

## Assigning object sides during training 

Next I used `rbinom()` to return a random sequence of 1s and 0s which I
converted to left and right which would determine which side of the tank an
object would end up on. To get the random seed I randomly picked numbers until I
got one which produced equal amounts of 1s and 0s.

Then I ran a series of dplyr calls which creates a new variable called `trial`
that counts from 1 to `nrow()` of the data frame separated by id. There are 20
levels of `trial`. The object sides for trial 0 was obtained by reversing the
object sides of trial 1. 

```{r treatment-assignment, echo=FALSE}
# Creating IDs
id <- as.factor(c(
  "1a", "1b", "2a", "2b", "3a", "3b", "4a", "4b",
  "5a", "5b", "6a", "6b", "7a", "7b", "8a", "8b"
))

# Get treatment assignments
set.seed(123)
treatment <- complete_ra(
  N = 16,
  num_arms = 2,
  conditions = c("green-trained", "blue-trained")
)

set.seed(123)
object_pair <- complete_ra(
  N = 20,
  num_arms = 4,
  conditions = c("pair-1", "pair-2", "pair-3", "pair-4")
)

# Get object sides for training
set.seed(52)
object_sides_training <- rbinom(20, 1, 0.5)

# Create a data frame with trials and object sides
data_sheet <- data.frame(id = rep(id, times = 20)) %>%
  group_by(id) %>%
  mutate(trial = 1:n()) %>%
  ungroup() %>%
  mutate(
    object.side =
      ifelse(
        trial %% 2 == 1,
        object_sides_training, 1 - object_sides_training
      )
  )

# Change object side to factor and re-code from numbers to words
data_sheet$object.side <- as.factor(data_sheet$object.side)
data_sheet$object.side <- recode_factor(data_sheet$object.side,
  "1" = "left",
  "0" = "right"
)

# Assigning object pairs to trials
data_sheet$trial <- as.factor(data_sheet$trial) # convert trial to factor
names(object_pair) <- levels(data_sheet$trial) # get the list of trial numbers to associate object pairs to
data_sheet$object_pair <- object_pair[as.character(data_sheet$trial)]

# Add individual treatments to data sheet
data_sheet <- cbind(data_sheet, treatment)
```

```{r, echo=FALSE, include=FALSE, eval=FALSE}
data_sheet %>% knitr::kable()
```

## Assigning object sides during tests

Trial 21 had its object sides obtained with the following script

```{r}
set.seed(55461)
test_sides <- complete_ra(
  N = 16,
  num_arms = 2,
  conditions = c("left", "right")
)

test_sides %>% kable()
```

Trial 23 had its object sides obtained by reversing the object sides of trial 21
and trial 25 had its object sides obtained by reversing the object sides of
trial 23.

Trial 27 had its object sides obtained with the following script and trial 29
had its object sides obtained by reversing the object sides of trial 27.

```{r}
set.seed(71)
test_sides <- complete_ra(
  N = 16,
  num_arms = 2,
  conditions = c("left", "right")
)

test_sides %>% kable()
```

Refresher trials 22 and 24 had their object sides obtained with the following
script. Refresher trials 26 and 28 had their object sides obtained by reversing
trials 22 and 24 respectively.

```{r}
# Get object sides for training
set.seed(52)
object_sides_refresher <- rbinom(2, 1, 0.5)

# Create a data frame with trials and object sides
data_sheet_refresher <- data.frame(id = rep(id, times = 2)) %>%
  group_by(id) %>%
  mutate(trial = 1:n()) %>%
  ungroup() %>%
  mutate(
    object.side =
      ifelse(
        trial %% 2 == 1,
        object_sides_refresher, 1 - object_sides_refresher
      )
  )

# Change object side to factor and re-code from numbers to words
data_sheet_refresher$object.side <- as.factor(data_sheet_refresher$object.side)
data_sheet_refresher$object.side <- recode_factor(data_sheet_refresher$object.side,
  "1" = "left",
  "0" = "right"
)
```

```{r, echo=FALSE}
data_sheet_refresher %>% knitr::kable()
```


